---
title: "Homework 3"
author: "Piotr Pasza Storozenko"
date: "29 03 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data
```{r message=FALSE, warning=FALSE}
library(haven)
library(mlr)
library(tidyverse)
```

```{r}
# Download data from https://stacks.stanford.edu/file/druid:hg921sg6829/HCMST_2017_public_data_v1.1_stata.zip and extract to the same folder as raport.
data <- read_stata('HCMST 2017 fresh sample for public sharing draft v1.1.dta')
```

## Choosed features

* `ppethm` - Race /Etnicity
* `age_when_met` - Self explaining
* `Q27` - Did you and your partner grow up in the same city or town?
* `w6_sex_frequency` - Frequency of heaving sex with partner during last 12 months.

```{r}
raw_df <- data %>% 
    select(c(ppethm, age_when_met, Q27, w6_sex_frequency, ppage, S1)) %>% 
    na.omit() 
labels <- sapply(1:ncol(raw_df), function(i) attributes(raw_df[[i]])$label)
legends <- sapply(1:ncol(raw_df), function(i) {
    labs <- attributes(raw_df[[i]])$labels
    labs[labs >0]
})
choosed_df <- raw_df %>% 
    mutate(ppethm = as.factor(ppethm)) %>% 
    filter(Q27 != -1) %>% 
    mutate(Q27 = as.factor(Q27)) %>% 
    filter(ppage > 0) %>% 
    mutate(ppage = as.numeric(ppage)) %>% 
    mutate(S1 = as.factor(S1)) %>% 
    filter(w6_sex_frequency != -1) %>% 
    mutate(w6_sex_frequency = as.numeric(w6_sex_frequency))
```

```{r warning=FALSE}
task <- makeClassifTask('hc',choosed_df, 'S1')
lrnr <- makeLearner('classif.randomForest', predict.type = 'prob')
model <- train(lrnr, task)
```

## Breakdown

The described person is White, Non-Hispanic, met with a partner at 37.
They grew up in different cities.
They do have sex once or twice a week.
Now this person is 52 and is married.

```{r fig.height=9, warning=FALSE}
obs <- choosed_df[1337,]
res <- predict(model, newdata = choosed_df)$data$prob.1 %>% mean()

df <- choosed_df
n <- nrow(df)
for (i in 1:5) {
  df[[i]] <- rep(obs[[i]], length.out = n)
  res[i+1] <- predict(model, newdata =  df)$data$prob.1 %>% mean()
}
difs <- diff(res)
labels[[2]] <- "Age when met?"
labels[[6]] <- NULL
labs <- unlist(labels)
labs <- c("Intercept", labs)
dta <- data.frame(Feature=as.character(labs), prob=res)

dta$end <- cumsum(c(0,difs)) + res[1]
dta$start <- c(0, head(dta$end, -1))
dta$id <- 1:6
dta$ud <- c(0,1,-1,0,0,1) %>% as.factor()
dta %>% ggplot(aes(rob, fill = ud)) + 
  geom_rect(aes(x=Feature, xmin = id - 0.45, xmax = id + 0.45, ymin = end, ymax = start)) +
  scale_x_discrete(labels = dta$Feature) +
  xlab("Feature") +
  ylab('Probability of being merried') +
  scale_fill_discrete("Trend", labels = c("Lose","None","Gain")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, fig.height=9, warning=FALSE}
obs <- choosed_df[1337,]
res <- predict(model, newdata = choosed_df)$data$prob.1 %>% mean()

df <- choosed_df
n <- nrow(df)
seqq <- c(5,3,1,2,4)
for (i in seqq) {
  df[[i]] <- rep(obs[[i]], length.out = n)
  res[i+1] <- predict(model, newdata =  df)$data$prob.1 %>% mean()
}
difs <- diff(res)
labels[[2]] <- "Age when met?"
labels[[6]] <- NULL
labs <- unlist(labels)
labs <- c("Intercept", labs[seqq])
dta <- data.frame(Feature=labs, prob=res)

dta$end <- cumsum(c(0,difs)) + res[1]
dta$start <- c(0, head(dta$end, -1))
dta$id <- 1:6
dta$ud <- c(0,1,-1,1,1,-1) %>% as.factor()
dta %>% ggplot(aes(prob, fill = ud)) + geom_rect(aes(x = Feature,
     xmin = id - 0.45, xmax = id + 0.45, 
     ymin = end,
     ymax = start))+
  xlab("Feature") +
  scale_x_discrete(labels = dta$Feature) +
  ylab('Probability of being merried') +
  scale_fill_discrete("Trend", labels = c("Lose","None","Gain")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Small summary

Effects from both orders differs a bit.
Only Age remains as the main factor, that boosts the probability of being married.