---
title: "Homework 1"
author: "Piotr Pasza Storozenko"
date: "1 03 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ceteris Paribus

## Read data
```{r message=FALSE, warning=FALSE}
library(haven)
library(mlr)
library(tidyverse)
```

```{r}
# Download data from https://stacks.stanford.edu/file/druid:hg921sg6829/HCMST_2017_public_data_v1.1_stata.zip and extract to the same folder as raport.
data <- read_stata('HCMST 2017 fresh sample for public sharing draft v1.1.dta')
```

## Choosed features

* `ppethm` - Race /Etnicity
* `age_when_met` - Self explaining
* `Q27` - Did you and your partner grow up in the same city or town?
* `w6_sex_frequency` - Frequency of heaving sex with partner during last 12 months.

```{r}
raw_df <- data %>% 
    select(c(ppethm, age_when_met, Q27, w6_sex_frequency, ppage, S1)) %>% 
    na.omit() 
labels <- sapply(1:ncol(raw_df), function(i) attributes(raw_df[[i]])$label)
legends <- sapply(1:ncol(raw_df), function(i) {
    labs <- attributes(raw_df[[i]])$labels
    labs[labs >0]
})
choosed_df <- raw_df %>% 
    mutate(ppethm = as.factor(ppethm)) %>% 
    filter(Q27 != -1) %>% 
    mutate(Q27 = as.factor(Q27)) %>% 
    filter(ppage > 0) %>% 
    mutate(ppage = as.numeric(ppage)) %>% 
    mutate(S1 = as.factor(S1)) %>% 
    filter(w6_sex_frequency != -1) %>% 
    mutate(w6_sex_frequency = as.numeric(w6_sex_frequency))
```

```{r warning=FALSE}
task <- makeClassifTask('hc',choosed_df, 'S1')
lrnr <- makeLearner('classif.randomForest', predict.type = 'prob')
model <- train(lrnr, task)
# rdesc = makeResampleDesc("Subsample", iters = 10, split = 3/4)
# benchmark(lrnr, task, rdesc, measures = list(f1))
```

```{r}
repr_inds <- sample(nrow(choosed_df), 100)

uni_vals <- lapply(1:(ncol(choosed_df)-1), function(c) {
    tmp <- unique(choosed_df[,c])
    tmp[order(tmp[[1]]),]
})

r1 <- sapply(repr_inds, function(ind){
    obs <- choosed_df[ind,]
    lapply(1:length(uni_vals), function(f){
        predict(model, newdata = cbind(uni_vals[[f]], obs[-f]))$data[,2]
    })
})
```

```{r}
r2 <- sapply(1:nrow(r1), function(i){
    do.call(cbind, r1[i,]) %>% as.data.frame() %>% gather(key = 'Person') %>% cbind(feature = uni_vals[[i]])
})
```

```{r}
plots <- lapply(1:ncol(r2), function(k){
    rdf <- r2[,k] %>% as.data.frame()
    colnames(rdf) <- c('Person', 'value', 'feature')
    if (k == 4) {
      rdf <- rdf %>% mutate(feature = as.factor(feature))
    }
    p <- rdf %>% 
        ggplot(aes(x = feature, y= value, group = Person)) +
        geom_line(color='grey') +
        xlab(labels[k]) +
        ylab('Probability of being merried') +
        scale_x_discrete(labels = as.character(names(legends[[k]])))
    
    mean_valdf <- rdf %>% 
        group_by(feature) %>% 
        summarise(mean = mean(value)) %>% 
        cbind(Person=101)
    p + geom_line(data = mean_valdf, aes(feature, mean))  +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
})
```

## Plots

```{r echo=FALSE}
plots[[1]]
```

```{r echo=FALSE, message=FALSE}
plots[[2]] + scale_x_continuous() + xlab("Age when met?")
```

```{r echo=FALSE}
plots[[3]]
```

```{r echo=FALSE}
plots[[4]]
```


```{r echo=FALSE, message=FALSE}
plots[[5]] + scale_x_continuous() 
```

## Bonus - histograms
```{r echo=FALSE}
choosed_df %>% ggplot(aes(x=ppethm)) +
  geom_bar() + 
  xlab(labels[1]) + 
  scale_x_discrete(labels = names(legends[[1]])) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
choosed_df %>% ggplot(aes(x=age_when_met)) +
  geom_histogram()+ 
  xlab("Age when met?") + 
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r echo=FALSE}
choosed_df %>% ggplot(aes(x=Q27)) +
  geom_bar()+ 
  xlab(labels[3]) + 
  scale_x_discrete(labels = names(legends[[3]])) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r echo=FALSE}
choosed_df %>% 
  mutate(w6_sex_frequency = as.factor(w6_sex_frequency)) %>% 
  ggplot(aes(x=w6_sex_frequency)) +
  geom_bar()+ 
  xlab(labels[4]) + 
  scale_x_discrete(labels = names(legends[[4]])) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
choosed_df %>% ggplot(aes(x=ppage)) +
  geom_histogram()+ 
  xlab(labels[5]) + 
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
summary(choosed_df$ppage)
```


## Funny fact

Model shows that heaving sex less frequently is related to being married.
It's not model speccific fact, SVMs and logistic regression shows the same.
I'm not an expert but I think this can be related to the fact that 1/4 of asked people are older then 61.