---
title: "Homework 4"
author: "Piotr Pasza Storozenko"
date: "09 04 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data
```{r message=FALSE, warning=FALSE}
library(haven)
library(mlr)
library(plyr)
library(tidyverse)
library(cluster)
```

```{r}
# Download data from https://stacks.stanford.edu/file/druid:hg921sg6829/HCMST_2017_public_data_v1.1_stata.zip and extract to the same folder as raport.
data <- read_stata('HCMST 2017 fresh sample for public sharing draft v1.1.dta')
```

## Choosed features

* `ppethm` - Race /Etnicity
* `age_when_met` - Self explaining
* `Q27` - Did you and your partner grow up in the same city or town?
* `w6_sex_frequency` - Frequency of heaving sex with partner during last 12 months.

```{r}
set.seed(1337)
raw_df <- data %>% 
    select(c(ppethm, age_when_met, Q27, w6_sex_frequency, ppage, S1)) %>% 
    na.omit() 
labels <- sapply(1:ncol(raw_df), function(i) attributes(raw_df[[i]])$label)
legends <- sapply(1:ncol(raw_df), function(i) {
    labs <- attributes(raw_df[[i]])$labels
    labs[labs >0]
})
choosed_df <- raw_df %>% 
    mutate(ppethm = as.factor(ppethm)) %>% 
    filter(Q27 != -1) %>% 
    mutate(Q27 = as.factor(Q27)) %>% 
    filter(ppage > 0) %>% 
    mutate(ppage = as.numeric(ppage)) %>% 
    mutate(S1 = as.factor(S1)) %>% 
    filter(w6_sex_frequency != -1) %>% 
    mutate(w6_sex_frequency = as.numeric(w6_sex_frequency))
```

```{r warning=FALSE}
task <- makeClassifTask('hc',choosed_df, 'S1')
lrnr <- makeLearner('classif.randomForest', predict.type = 'prob')
model <- train(lrnr, task)
```

## Lime

The described person is White, Non-Hispanic, met with a partner at 37.
They grew up in different cities.
They do have sex once or twice a week.
Now this person is 52 and is married.

Task:

Let's see how the model behaves around this person. Assume that we're interested only in White, Non-Hispanic, that grew up in a different city than their partner.

```{r fig.height=9, warning=FALSE}
obs <- choosed_df[1337,]

ages <- rnorm(200, 37, 10)
ppages <- rnorm(200, 52, 10)
sexfs <- choosed_df$w6_sex_frequency %>% unique()

df_agu <- expand.grid(ppethm = as.factor(1), age_when_met = ages,Q27=as.factor(1), w6_sex_frequency=sexfs,ppage =ppages) %>% as.data.frame()
levels(df_agu$ppethm) <- 1:5
levels(df_agu$Q27) <- 1:2
rand_inds <- sample(nrow(df_agu), 2000)
df_agu_t <- df_agu[rand_inds,] %>% as_tibble()

pertub_pred <- predict(model, newdata = df_agu_t)
pertub_merried <- pertub_pred$data$response
df_agu_t$S1 <- pertub_merried

df_pert <- rbind(obs,df_agu_t)
dists <- daisy(df_pert, metric = 'gower')
dists_m <- as.matrix(dists)
dists_from_red <- dists_m[1,2:ncol(dists_m)]
df_agu_train <- df_agu_t %>% select(age_when_met, w6_sex_frequency, ppage, S1)

task2 <- makeClassifTask('hc2',df_agu_train, 'S1')
lrnr2 <- makeLearner('classif.binomial', predict.type = 'prob')
model2 <- train(lrnr2, task2, weights = dists_from_red)

pred <-  predict(model2, newdata = df_agu_train)
```

Logistic regression model has a decent performance:
```{r}
performance(pred, measures = list(acc, f1))
```

Model params:
```{r}
getLearnerModel(model2)
```


## Small summary

Compared to the previous homework, age (variable ppage) is not such a big factor.
More important is the age when met and sex frequency.