---
title: "Homework 2"
author: "Piotr Pasza Storozenko"
date: "23 03 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Read data
```{r message=FALSE, warning=FALSE}
library(haven)
library(mlr)
library(tidyverse)
```

```{r}
# Download data from https://stacks.stanford.edu/file/druid:hg921sg6829/HCMST_2017_public_data_v1.1_stata.zip and extract to the same folder as raport.
data <- read_stata('HCMST 2017 fresh sample for public sharing draft v1.1.dta')
```

## Choosed features

* `ppethm` - Race /Etnicity
* `age_when_met` - Self explaining
* `Q27` - Did you and your partner grow up in the same city or town?
* `w6_sex_frequency` - Frequency of heaving sex with partner during last 12 months.

```{r}
raw_df <- data %>% 
    select(c(ppethm, age_when_met, Q27, w6_sex_frequency, ppage, S1)) %>% 
    na.omit() 
labels <- sapply(1:ncol(raw_df), function(i) attributes(raw_df[[i]])$label)
legends <- sapply(1:ncol(raw_df), function(i) {
    labs <- attributes(raw_df[[i]])$labels
    labs[labs >0]
})
choosed_df <- raw_df %>% 
    mutate(ppethm = as.factor(ppethm)) %>% 
    filter(Q27 != -1) %>% 
    mutate(Q27 = as.factor(Q27)) %>% 
    filter(ppage > 0) %>% 
    mutate(ppage = as.numeric(ppage)) %>% 
    mutate(S1 = as.factor(S1)) %>% 
    filter(w6_sex_frequency != -1) %>% 
    mutate(w6_sex_frequency = as.numeric(w6_sex_frequency))
```

```{r warning=FALSE}
task <- makeClassifTask('hc',choosed_df, 'S1')
lrnr <- makeLearner('classif.randomForest', predict.type = 'prob')
model <- train(lrnr, task)
```

## Local Conditional Expectations

```{r}
repr_inds <- sample(nrow(choosed_df), 100)

uni_vals <- lapply(1:(ncol(choosed_df)-1), function(c) {
    tmp <- unique(choosed_df[,c])
    tmp[order(tmp[[1]]),]
})

r1 <- sapply(repr_inds, function(ind){
    obs <- choosed_df[ind,]
    lapply(1:length(uni_vals), function(f){
        cbind(PersonID = ind, cbind(uni_vals[[f]], obs[-f]), proba = predict(model, newdata = cbind(uni_vals[[f]], obs[-f]))$data[,2])
    })
})
```

```{r}
r2 <- lapply(1:nrow(r1), function(i){
    do.call(rbind, r1[i,])
})
```


```{r}
d2 <- r2[[2]] %>% as_tibble()

mean_val <- d2 %>% group_by(age_when_met, ppethm) %>% summarise(mean=mean(proba))

d2 %>% ggplot(aes(x=age_when_met, y=proba, group = PersonID)) +
  geom_line(color='grey') +
  geom_line(data = mean_val, aes(age_when_met, mean, group=ppethm, color = ppethm)) +
  scale_color_discrete(labels[[1]],labels=names(legends[[1]])) +
  xlab("Age when met?") +
  ylab('Probability of being merried') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
d4 <- r2[[4]] %>% as_tibble()
mean_val <- d4 %>% group_by(w6_sex_frequency, ppethm) %>% summarise(mean=mean(proba))

d4 %>% ggplot(aes(x=w6_sex_frequency, y=proba, group = PersonID)) +
  geom_line(color='grey') +
  geom_line(data = mean_val, aes(w6_sex_frequency, mean, group=ppethm, color = ppethm)) +
  scale_color_discrete(labels[[1]],labels=names(legends[[1]])) +
  xlab(labels[[4]]) +
  ylab('Probability of being merried') +
  scale_x_continuous(labels = as.character(names(legends[[4]]))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
d1 <- r2[[1]] %>% as_tibble()
d1 <- d1 %>% mutate(w6_sex_frequency = w6_sex_frequency %>% as.factor())
mean_val <- d1 %>% group_by(ppethm, w6_sex_frequency) %>% summarise(mean=mean(proba))

name <- "During the last 12 months,\nabout how often\ndid you have sex with\n[Partner name]?"

d1 %>% ggplot(aes(x=ppethm, y=proba, group = PersonID)) +
  geom_line(color='grey') +
  geom_line(data = mean_val, aes(ppethm, mean, group=w6_sex_frequency, color = w6_sex_frequency))  +
  scale_x_discrete(labels = as.character(names(legends[[1]]))) +
  scale_color_discrete(name,labels=names(legends[[4]])) +
  ylab('Probability of being merried') +
  theme_minimal() +
  xlab(labels[[1]]) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Small summary
The first plot shows that White, Non-Hispanic people, generally speaking, are meeting their partners earlier then Hispanic.
A shape of curve related to "2+ Races" can heavily be biased by a low amount of people of this race in the dataset.

Second and third plot shows, that there are no strong interactions between the amount of sex in a relationship and the race.
