---
title: "Homework 6"
author: "Piotr Pasza Storozenko"
date: "16 05 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data
```{r message=FALSE, warning=FALSE}
library(haven)
library(mlr)
library(tidyverse)
```

```{r}
# Download data from https://stacks.stanford.edu/file/druid:hg921sg6829/HCMST_2017_public_data_v1.1_stata.zip and extract to the same folder as raport.
data <- read_stata('HCMST 2017 fresh sample for public sharing draft v1.1.dta')
```

## Choosed features

* `ppethm` - Race / Etnicity
* `ppage` - Current Age
* `age_when_met` - Self explaining
* `Q27` - Did you and your partner grow up in the same city or town?
* `w6_sex_frequency` - Frequency of heaving sex with partner during last 12 months.

```{r}
set.seed(1337)
raw_df <- data %>% 
    select(c(ppethm, age_when_met, Q27, w6_sex_frequency, ppage, S1)) %>% 
    na.omit() 
labels <- sapply(1:ncol(raw_df), function(i) attributes(raw_df[[i]])$label)
legends <- sapply(1:ncol(raw_df), function(i) {
    labs <- attributes(raw_df[[i]])$labels
    labs[labs >0]
})
choosed_df <- raw_df %>% 
    mutate(ppethm = as.factor(ppethm)) %>% 
    filter(Q27 != -1) %>% 
    mutate(Q27 = as.factor(Q27)) %>% 
    filter(ppage > 0) %>% 
    mutate(ppage = as.numeric(ppage)) %>% 
    mutate(S1 = as.factor(S1)) %>% 
    filter(w6_sex_frequency != -1) %>% 
    mutate(w6_sex_frequency = as.numeric(w6_sex_frequency)) %>% 
    as.data.frame()
```

```{r warning=FALSE}
test_ind <- sample(1:nrow(choosed_df), 500)
train_df <- choosed_df[-test_ind,]
test_df <- choosed_df[test_ind,]

task <- makeClassifTask('hc',train_df, 'S1')
lrnr <- makeLearner('classif.ranger', predict.type = 'prob')
```

```{r}
model <- train(lrnr, task)
```


```{r}
test_df_ext <- test_df
test_df_ext$prediction <- getPredictionResponse(predict(model, newdata = test_df)) %>% as.numeric()
test_df_ext$S1 <- test_df_ext$S1 %>% as.numeric()
test_df_ext <- test_df_ext %>% mutate(difference = S1 - prediction) 
```

```{r message=FALSE, warning=FALSE}
test_df_ext %>% ggplot(aes(age_when_met, difference)) +
  geom_point() +
  geom_smooth() +
  xlab("Age when met") +
  ylab("Error y - y_hat") +
  theme_minimal()

test_df_ext %>% ggplot(aes(w6_sex_frequency, difference)) +
  geom_point() +
  geom_smooth() +
  xlab("Sex frequency") +
  ylab("Error y - y_hat")+
  theme_minimal()
  
test_df_ext %>% ggplot(aes(S1, difference)) +
  geom_point() +
  geom_smooth() +
  xlab("Merried?") +
  ylab("Error y - y_hat")+
  theme_minimal()
```


This three plots shows that model behaves _good_ witch respect to features `Age when met` and `Sex frequency` but it's generaly wrong more often when person is merried.

```{r, cache=TRUE}
y_pred <- getPredictionProbabilities(predict(model, newdata = train_df))
res <- sapply(1:nrow(train_df), function(i){
    ndata_t <- train_df[-i,]
    ntask <- makeClassifTask("hc2", data = ndata_t, target = "S1",
                                      positive = '1')
    learner_l <- makeLearner("classif.ranger", predict.type = 'prob')
    model_l <- train(learner_l, ntask)
    nprediction <- getPredictionProbabilities(predict(model_l, newdata = train_df))
    r <- mean((nprediction - y_pred)**2)
    r
})
```

```{r}
rdf_df <- data.frame(res = res, ind = 1:nrow(train_df))
rdf_df %>% ggplot(aes(ind, res)) +
    geom_smooth() + 
    geom_line() +
    geom_smooth() +
    ylab("Mean square error of prediction") +
    xlab("Index of observation") + 
    theme_minimal()

rdf_df %>% ggplot(aes(res)) + 
  geom_histogram() + 
  xlab("Mean square error") +
  ylab("Count") + 
  theme_minimal()
```

Above plots shows that there are no observations that are very heavily influencing.

The most influencing observation is the following:

```{r}
train_df[which.max(res),]                     
```


