---
title: "Homework 5"
author: "Piotr Pasza Storozenko"
date: "09 04 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data
```{r message=FALSE, warning=FALSE}
library(haven)
library(mlr)
library(plyr)
library(tidyverse)
library(cluster)
```

```{r}
# Download data from https://stacks.stanford.edu/file/druid:hg921sg6829/HCMST_2017_public_data_v1.1_stata.zip and extract to the same folder as raport.
data <- read_stata('HCMST 2017 fresh sample for public sharing draft v1.1.dta')
```

## Choosed features

* `ppethm` - Race / Etnicity
* `ppage` - Current Age
* `age_when_met` - Self explaining
* `Q27` - Did you and your partner grow up in the same city or town?
* `w6_sex_frequency` - Frequency of heaving sex with partner during last 12 months.

```{r}
set.seed(1337)
raw_df <- data %>% 
    select(c(ppethm, age_when_met, Q27, w6_sex_frequency, ppage, S1)) %>% 
    na.omit() 
labels <- sapply(1:ncol(raw_df), function(i) attributes(raw_df[[i]])$label)
legends <- sapply(1:ncol(raw_df), function(i) {
    labs <- attributes(raw_df[[i]])$labels
    labs[labs >0]
})
choosed_df <- raw_df %>% 
    mutate(ppethm = as.factor(ppethm)) %>% 
    filter(Q27 != -1) %>% 
    mutate(Q27 = as.factor(Q27)) %>% 
    filter(ppage > 0) %>% 
    mutate(ppage = as.numeric(ppage)) %>% 
    mutate(S1 = as.factor(S1)) %>% 
    filter(w6_sex_frequency != -1) %>% 
    mutate(w6_sex_frequency = as.numeric(w6_sex_frequency)) %>% 
    as.data.frame()
```

```{r warning=FALSE}
test_ind <- sample(1:nrow(choosed_df), 500)
train_df <- choosed_df[-test_ind,]
test_df <- choosed_df[test_ind,]

task <- makeClassifTask('hc',train_df, 'S1')
```

## Loss on perturbations RandomForest vs QDA

```{r}
learner1 <-  makeLearner("classif.randomForest", predict.type = 'prob')
model1 <- train(learner1, task)

learner2 <-  makeLearner("classif.qda", predict.type = 'prob')
model2 <- train(learner2, task)

y_pred <- predict(model1, newdata=test_df)
acc01 <- performance(y_pred, acc)
y_pred <- predict(model2, newdata=test_df)
acc02 <- performance(y_pred, acc)
acc0 <- mean(acc01, acc02)

r <- sapply(1:(ncol(choosed_df)-1), function(i) {
    act_df <- test_df
    act_df[,i] <- sample(act_df[,i])
    y_pred1 <- predict(model1, newdata=act_df)
    y_pred2 <- predict(model2, newdata=act_df)
    c(rf = performance(y_pred1, acc) - acc01,
      qda = performance(y_pred2, acc) - acc02,
      feature = colnames(test_df)[i])
})
t(r) %>% as_tibble() %>% 
  mutate_at(vars(rf.acc, qda.acc), as.numeric) %>% 
  gather(key = 'classif', value = 'acc', -feature) %>% 
  ggplot(aes(x=feature, y = acc, fill=classif)) +
  geom_col(position = 'dodge') +
  theme_minimal() +
  ylab("Acc difference after perturbation")
```

The biggest difference is on feature `ppage`. 
Mean acc for rf is `r acc01` and for qda is `r acc02`.

## PDP plot on `ppage` feature

```{r}
ages <- unique(test_df$ppage)
pdp <- sapply(ages, function(age){
    act_df <- test_df
    act_df$ppage <- age
    
    yp1 <- predict(model1, newdata = act_df)
    yp2 <- predict(model2, newdata = act_df)
    
    p1 <- getPredictionProbabilities(yp1)
    p2 <- getPredictionProbabilities(yp2)
    c(mean(p1), mean(p2))
})
```

```{r}
df_pdp <- cbind(data.frame(ages), t(pdp))
colnames(df_pdp) <- c('ppage', 'rf', 'qda')
df_pdpf <- df_pdp %>% gather("classif", "proba", -ppage)

df_pdpf %>% ggplot(aes(x=ppage, y=proba, color=classif)) +
  geom_line() +
  theme_minimal() +
  xlab("Participant age") +
  ylab("Probability of being married")
```

The PDP plot shows that RF model predictions are much more rapid change with respect to `ppage`, however, both models have similar general shape.
This can be caused by the simplicity of QDA model.